---
title: "Exam 1"
subtitle: "MATH 362"
author: "YOUR NAME HERE"
date: "due Saturday, October 8, 11:59pm"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
    theme: "paper"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

**Note: you may need to install one or more of the following packages before being able to knit. You can install the `kableExtra` package, for example, by running `install.packages("kableExtra")` in your console.**

```{r, message = FALSE, warning = FALSE}
library(infer)
library(kableExtra)
library(tidyverse)
```


# Problem 1

```{r prob-1}
data2019 <- c(3, 2, 7, 4, 0, 2, 2, 4, 6, 0,
       2, 8, 3, 3, 1, 1, 0, 2, 0, 2,
       4, 0, 1, 1, 2, 8, 1, 1, 4, 2, 
       4, 4, 1, 1, 2, 1, 5, 3, 3, 1, 
       1, 12, 2, 0, 5, 8, 3, 9, 2, 2)

data2020 <- c(0, 4, 4, 3, 8, 2, 5, 5, 2, 3, 
       3, 4, 1, 3, 4, 1, 3, 11, 5, 2,  
       1, 2, 7, 4, 4, 4, 8, 2, 1, 4,  
       7, 4, 0, 1, 4, 8, 4, 3, 5, 11,  
       9, 3, 2, 2, 4, 3, 0, 0, 12, 4)
```

a)

b)

c)

d)

e)

f)

g)

h)


# Problem 2

### Part a

```{r prob-2}

```

> YOUR ANSWER HERE

### Part b

> YOUR ANSWER HERE

# Problem 3

```{r}
#read in population data from Google sheets URL
population <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vQ7Pn9SuN6KzdvxgqMG98iR8J3gorfkPLCVbF9VV4zwna5eY6-21M9ZV98MmxpX6LQ1iN8ChZur1rXe/pub?gid=429421419&single=true&output=csv", 
                       show_col_types = FALSE)

#draws 10,000 random samples
samples <- population %>% 
  rep_sample_n(size = 100, reps = 10000)

#computes sample proportion in each sample
sample_props <- samples %>% 
  group_by(replicate) %>% 
  summarize(phat = sum(smoke == "Yes")/n())
```

### Part a

```{r prob-3a}

```

> YOUR ANSWER HERE

### Part b

> YOUR ANSWER HERE

### Part c

```{r prob-3c}

```

### Part d

> YOUR ANSWER HERE

# Problem 4

A) 

B) 

C)

D)

E)

# Problem 5

A) 

B)

C)

D)

E)
