---
title: "Problem Set 04"
subtite: "TEMPLATE"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
    theme: "cosmo"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Summary / Reflection Prompts

*RP1: What were the main concepts covered in this assignment?*

*RP2: What's one thing you understand better after completing these problems?*

*RP3: What problems gave you the most trouble? What was difficult about them/where did you get stuck?*

## Exercise 1

Text...

```{r ex-1}

```

Text...

## Exercise 2

```{r ex-2}

```

Text...

## Exercise 3

```{r ex-3}

```

Text...

## Exercise 4

```{r ex-4}

```

Text...

## Exercise 5

```{r ex-5}

```

Text...

## Exercise 6

Text...

## Exercise 7

Text... 
Sample LaTeX syntax: $E(\overline{X})=\mu$

## Exercise 8

Text...

## Exercise 9

```{r ex-9, eval = FALSE}
#psuedo-code here

```


## Exercise 10

Text...

## Exercise 11

```{r ex-11}

```

Text...

## Exercise 12

```{r ex-12}

```

Text...

## Exercise 13

```{r ex-13}

```

Text...

## Exercise 14

Text...

## Exercise 15

Text...

## Exercise 16

```{r ex-16}

```

Text...

## Exercise 17

Text...

## Exercise 18

```{r ex-18}

```

Text...

## Exercise 19

```{r ex-19}

```

Text...

## Exercise 20

Text...

## Exercise 21

Text...