---
title: "Problem Set 06"
subtite: "TEMPLATE"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
    theme: "cosmo"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r}
library(NHANES)
library(infer)
library(tidyverse)
```


## Summary / Reflection Prompts

*RP1: What were the main concepts covered in this assignment?*

*RP2: What's one thing you understand better after completing these problems?*

*RP3: What problems gave you the most trouble? What was difficult about them/where did you get stuck?*

## Exercise 1

## Exercise 2

## Exercise 3

## Exercise 4

## Exercise 5

## Exercise 6

## Exercise 7


# Simulation exploration

```{r, message = FALSE, warning = FALSE}
adults <- NHANES %>% 
  filter(Age >= 20)

p <- adults %>% 
  summarize(p = sum(Smoke100 == "Yes")/n()) %>% 
  pull
p

n <- 100
samples <- rep_sample_n(adults, n, 
                        replace = TRUE, 
                        reps = 10000)

sample_props <- samples %>% 
  group_by(replicate) %>% 
  summarize(phat = sum(Smoke100 == "Yes")/n()) 
```

## Exercise 8

The distribution appears normal (unimodal, symmetric), centered around 0.44 as expected. 

```{r}
ggplot(sample_props, aes(x = phat)) +
  geom_histogram(color = "white")
```

## Exercise 9

0.05. We know $H_0$ to be true, so $P(|Z| > 1.96) = 0.05$. $|Z| > 1.96$ will lead to a decision of Reject $H_0$, which is an incorrect decision in this case.  

## Exercise 10

```{r}
p0 <- 0.44
sample_props <- sample_props %>% 
  mutate(Z = (phat - p0)/sqrt(p0*(1-p0)/n),
         reject = if_else(abs(Z) > 1.96, 1, 0))

sample_props %>% 
  summarize(sum(reject)/n())
```

Even when $H_0$ is true, about 5% of samples will produce a $\hat{p}$ that are far enough from $p$ that we will incorrectly reject the null hypothesis. 

## Exercise 11

By re-arranging $$Z = \frac{\hat{p} - p_0}{\sqrt{\frac{p_0(1-p_0)}{n}}}$$ we get $$\hat{p} = Z\sqrt{\frac{p_0(1-p_0)}{n}} + p_0.$$

We can plug in -1.96 and 1.96 for Z to find the cutoff values for the critical region, on the scale of $\hat{p$.

```{r}
Z <- qnorm(0.975)
SE_phat <- sqrt(p0*(1-p0)/n)
lb_phat <- -Z*SE_phat + p0
ub_phat <- Z*SE_phat + p0
c(lb_phat, ub_phat)

ggplot(sample_props, aes(x = phat)) +
  geom_histogram(color = "white") +
  geom_vline(xintercept = lb_phat, color = "blue") +
  geom_vline(xintercept = ub_phat, color = "blue")
```

About 5% of the simulated values fall between `r lb_phat` and `r ub_phat`.

## Exercise 12

We expect to get a p-value < 0.05 about 5% of the time when $H_0$ is true. The p-value quantifies the probability that the sample proportion will fall more extreme than the one observed, assuming $H_0$ is true. Because $H_0$ is true here, there should be just the 2.5% in each tail should have p-values < 0.05. 

## Exercise 13

```{r}
sample_props <- sample_props %>% 
  mutate(p_value = 2*pnorm(-abs(Z)),
         p_onesided = pnorm(Z, lower.tail = FALSE),
         reject_p = if_else(p_value < 0.05, 1, 0))

sample_props %>% 
  summarize(sum(reject_p)/n()) 
```

Again, about 5% of the samples led to incorrect rejections of $H_0$. The same exact samples that had $|Z| > 1.96$ have p-value < 0.05 because these are equivalent ways of deciding a hypothesis test. 

## Exercise 14

Type I error - rejecting $H_0$ when it is true (and we know it to be true in this case). If we wanted to reduce this Type I error rate we can change the threshold - push it farther out in the tails by making the Z-cutoff larger, or equivalently the p-value smaller.

## Exercise 15

About 1% would fall in the critical region (|Z|>2.576). This would lead to a Type I error rate of 1% instead of 5%. The appropriate p-value cutoff would then be 0.01. 

## Exercise 16

A Type II error is possible. Because the null hypothesis is false and should be rejected, so the danger is a false negative. 

## Exercise 17

Note: determining critical regions depends entirely on the hypothesized value under the null, not the actual value. 

```{r}
Z <- qnorm(0.975)
p0 <- 0.5
SE_phat <- sqrt(p0*(1-p0)/n)
lb_phat <- -Z*SE_phat + p0
ub_phat <- Z*SE_phat + p0
c(lb_phat, ub_phat)
```


## Exercise 18

The Type II error rate is ~22%. In real life, we obviously don't know the true 0.44, but we use this to determine best/worst case scenario if the true parameter is "this large/small". 

```{r}
pnorm(lb_phat, mean = 0.44, sd = sqrt(0.44*(1-0.44)/n)) + pnorm(ub_phat, mean = 0.44, sd = sqrt(0.44*(1-0.44)/n), lower.tail = FALSE)

sample_props %>% 
  summarize(sum(phat < lb_phat | phat > ub_phat)/n())
```

<!-- ## Exercise 19 -->

<!-- ```{r} -->
<!-- n <- 500 -->
<!-- samples500 <- rep_sample_n(adults, n,  -->
<!--                         replace = TRUE,  -->
<!--                         reps = 10000) -->

<!-- sample_props500 <- samples500 %>%  -->
<!--   group_by(replicate) %>%  -->
<!--   summarize(phat = sum(Smoke100 == "Yes")/n())  -->

<!-- p0 <- 0.44 -->
<!-- n <- 100 -->
<!-- Z <- qnorm(0.975) -->
<!-- SE_phat <- sqrt(p0*(1-p0)/n) -->
<!-- lb_phat <- -Z*SE_phat + p0 -->
<!-- ub_phat <- Z*SE_phat + p0 -->
<!-- c(lb_phat, ub_phat) -->

<!-- ggplot(sample_props500, aes(x = phat)) + -->
<!--   geom_histogram(color = "white") + -->
<!--   geom_vline(xintercept = lb_phat, color = "blue") + -->
<!--   geom_vline(xintercept = ub_phat, color = "blue") -->
<!-- ``` -->


<!-- ## Exercise 20 -->

<!-- ```{r} -->
<!-- p0 <- 0.44 -->
<!-- n <- 500 -->
<!-- Z <- qnorm(0.975) -->
<!-- SE_phat <- sqrt(p0*(1-p0)/n) -->
<!-- lb_phat <- -Z*SE_phat + p0 -->
<!-- ub_phat <- Z*SE_phat + p0 -->
<!-- c(lb_phat, ub_phat) -->
<!-- ``` -->

<!-- This cutoff values are closer to 0.44 because the larger sample size narrowed the overall sampling distribution. The Z cut-off values remain the same, because Z is a standardized distribution that already accounts for the effect of different sample sizes. -->

<!-- ## Exercise 21 -->

<!-- This is describing Type I error rate again. -->

<!-- ```{r} -->
<!-- sample_props500 <- sample_props500 %>%  -->
<!--   mutate(Z = (phat - p0)/sqrt(p0*(1-p0)/n), -->
<!--          reject = if_else(abs(Z) > 1.96, 1, 0)) -->

<!-- sample_props500 %>%  -->
<!--   summarize(sum(Z > 1.96 | Z < -1.96)/n())  -->
<!-- ``` -->


