---
title: "Problem Set 01"
author: "TEMPLATE"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
    theme: "cosmo"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Summary / Reflection Prompts

*RP1: What were the main concepts covered in this assignment?*

*RP2: What's one thing you understand better after completing these problems?*

*RP3: What problems gave you the most trouble? What was difficult about them/where did you get stuck?*

```{r load-pkgs}
#load necessary packages
library(tidyverse)
```

## Exercise 1

Text...

## Exercise 2

Text...

## Exercise 3

```{r ex3, eval = FALSE}
# pseudo-code

```

Text...

## Exercise 4

Text...

```{r ex4-sims}
# simulation code

```

```{r ex4-viz}
# visualizations

```


## Exercise 5

Text...


## Exercise 6

```{r ex6}
# calculate margin of error

# calculate confidence interval

```


## Exercise 7

Text...

```{r ex7}
#simulation code
```

Text...

## Exercise 8

Text...

## Exercise 9

```{r ex9}

```


## Exercise 10

Text...


## Exercise 11

```{r ex11-scatterplot}

```

*(You may want to do the following exercises with pen(cil) and paper, but you don't need to type up all the math here; just enough so someone else can follow your logic)*

## Exericse 12

Text...

## Exercise 13

Text...

## Exercise 14

Text... 

## Exercise 15

```{r ex15}
# graph probability as function of w

```

## Exercise 16

Text...

## Exercise 17

Text...




