---
title: "Problem Set 04"
subtite: "TEMPLATE"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
    theme: "cosmo"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Summary / Reflection Prompts

*RP1: What were the main concepts covered in this assignment?*

*RP2: What's one thing you understand better after completing these problems?*

*RP3: What problems gave you the most trouble? What was difficult about them/where did you get stuck?*

## Exercise 1



## Exercise 2



## Exercise 3

```{r ex-3}

```

a. The parameter of interest is... 
b. The point estimate is...
c. The confidence interval is... 

## Exercise 4

```{r ex-4}

```

a. The parameter of interest is... 
b. The point estimate is...
c. We are ___% confident that... 

## Exercise 5

```{r ex-5}

```

a. The parameter of interest is... 
b. The point estimate is...
c. We are ___% confident that... 

## Exercise 6

```{r ex-6}

```

a. The parameter of interest is... 
b. The point estimate is...
c. We are ___% confident that... 

## Exercise 7

```{r ex-7}

```

## Exercise 8



## Exercise 9




## Exercise 10



## Exercise 11




## Exercise 12



## Exercise 13



## Exercise 14



## Exercise 15



## Exercise 16



## Exercise 17


## Exercise 18


## Exercise 19


## Exercise 20


## Exercise 21


## Exercise 22


## Exercise 23


## Exercise 24



