---
title: "Problem Set 08"
subtitle: "Week 05 Day 01"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
    theme: "cosmo"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
```


# Learning objectives

# Cohen's d

In many different research contexts, the parameter of interest is a mean difference. For example we are interested in the difference between the mean in the treatment group $\mu_T$ and the mean in the control group $\mu_C$. It is common practice to standardize this difference by the population standard deviation. That is, the parameter of interest is $$\delta = \frac{\mu_T - \mu_C}{\sigma}$$. We estimate $\delta$ by $$d = \frac{\overline{Y}_T - \overline{Y}_C}{S_p^2}$$

Under some standard assumptions (that aren't too hard to meet in many research contexts), $d \sim N(0,1)$. The estimator $d$ above, is commonly known as Cohen's $d$, named after psychologist and statistician Jacob Cohen. 

When designing a study, you want to determine:

+ What is a practically meaningful value of $\delta$ in my research context?
+ What sample size do I need to be able to detect that effect?

Or alternatively, 

+ What is the largest sample size I can afford with my budget?
+ What is the minimal detectable effect size (MDES) with that sample size?

# Simulation exploration (in app)

Go to the following app: [https://rpsychologist.com/d3/nhst/](https://rpsychologist.com/d3/nhst/)

1. Assume that you want to detect an effect size of d = 0.1. How large of a sample do you need for the following scenarios:

    a. Type I error of 0.05 and power of 0.80
    b. Type I error of 0.01 and power of 0.80
    c. Type I error of 0.01 and power of 0.90
    
2. Let's say you want to power your test at 80% and restrict your Type I error to 0.01 and can only afford a sample size of 50. What's the smallest Cohen's d you will be able to detect under these constraints?

3. Let's say the value you determined in #2 is unrealistically large for your research context, and you really need to be able to detect an effect size of 0.3. However, you don't have access to any more money - $n = 50$ is still all that you can afford. 

    a. What could you change about your design to be able to detect 0.3? Come up with at least two alternative designs. 
    b. Comment on the tradeoffs in your designs.
    
4. Suppose you want to power your study at 80% and need to detect an effect of 0.5. What happens to the necessary sample size as alpha decreases? Describe why this happens. Describe the trade-off here in practical terms.

5. Suppose a Type I error is relatively risky in your context and therefore you set $\alpha = 0.01$. You need to detect an effect of 0.5. What happens to the necessary sample size as power decreases? Describe why this happens. Describe the trade-off here in practical terms.

6. Suppose you decide to power your study at 80% and are okay with a Type I error of 5%. What happens to the necessary sample size as $d$ changes? Describe why this happens. Describe the trade-off here in practical terms. 

7. Suppose you're serving as a statistical consultant and your client  says they have budgeted for a sample size of $n = 50$. They mention that in a pilot study they estimated their treatment improved outcomes by about d = 0.1. Explain how you would go about articulating to them that this research design is not set up for success. 

## Submission

Once you have completed all Exercises and written up your results in the Template .Rmd provided, take about 5-10 minutes to respond to the reflection prompts (RP) at the top of the Template:

+ (RP1): *What were the main concepts covered in this assignment?*

+ (RP2): *What's one thing you understand better after completing these problems?*

+ (RP3): *What problems gave you the most trouble? What was difficult about them/where did you get stuck?*

Knit your document one final time, then upload your .html file to Problem Set 05 on Canvas. 

## Grading (50pts)

| Component | Points |
|:----------|:-------|
| RP1      | 5      |
| RP2      | 5    |
| RP3      | 5    |
| Formatting & Clarity | 5 |
| Exercise completion | 30 |