---
title: "Problem Set 08"
subtitle: "Week 08"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
    theme: "cosmo"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


# Learning objectives


# Exercises

## By hand or typed

### Exercise 1

[HTZ: 6.5-1] Show that the residuals $Y_i = \hat{Y}_i, \ \ \ i= 1,2,...,n$ from the least squares regression model sum to 0.

### Exercise 2

[HTZ 6.5-2] In some situations where the regression model is useful, it is known that when $X = 0$ the mean of $Y$ is 0. In this case, the model $Y_i = \beta x_i + \epsilon_i,$ where $\epsilon_i$ $i = 1,2,...,n$ are independent with $\epsilon_i \sim N(0, \sigma^2)$.

a. Obtain the MLEs $\hat{\beta}$ and $\hat{\sigma}^2$ of $\beta$ and $\sigma^2$ under this model
b. Find the distribution of $\hat{\beta}$. 
c. Suppose a sample of size $n = 2$ consists of the observations $(x_1, y_1) = (1,1)$ and $(x_2, y_2) = (2,1)$. Calculate the estimates of $\beta$ and $\sigma^2$ from part (a) for these data, and plot the points and fitted line on the same graph
d. For the data in part (c), let $\hat{y_i} = \hat{\beta}x_i$ for $i = 1,2$. Calculate the residuals $y_i - \hat{y}_i$ and show that they do not sum to 0. 

### Exercise 3

Under the assumptions of the simple linear regression model, prove that a $(1 - \alpha)100\%$ confidence interval for the intercept parameter $\alpha$ is given by:

$$a \pm t_{\alpha/2}\left(\sqrt{\frac{\hat{\sigma}^2}{n-2}}\right)$$
### Exercise 4

```{r, echo = FALSE, eval = FALSE}
library(tidyverse)

senic <- read.table("/Volumes/GoogleDrive/My Drive/STAT 511/homework/data/APPENC01.txt", quote="\"", comment.char="")

colnames(senic) <- c("ID", "length", "age", "infection_risk",
                     "ratio_culture", "ratio_xray", "num_beds",
                     "med_school", "region", "avg_census",
                     "num_nurses", "avail_facilities")

write_csv(senic, "./PROBLEM_SETS/data/senic.csv")
```


## Submission

Once you have completed all Exercises and written up your results in the Template .Rmd provided, take about 5-10 minutes to respond to the reflection prompts (RP) at the top of the Template:

+ (RP1): *What were the main concepts covered in this assignment?*

+ (RP2): *What's one thing you understand better after completing these problems?*

+ (RP3): *What problems gave you the most trouble? What was difficult about them/where did you get stuck?*

Knit your document one final time, then upload your .html file to Problem Set 07 on Canvas. 

## Grading (50pts)

| Component | Points |
|:----------|:-------|
| RP1      | 5      |
| RP2      | 5    |
| RP3      | 5    |
| Formatting & Clarity | 5 |
| Exercise completion | 30 |