---
title: "Problem Set 06"
subtitle: "Week 05"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
    theme: "cosmo"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r, include = FALSE}
library(tidyverse)
library(NHANES)
library(cowplot)
library(infer)
```


# Learning objectives

+ Conduct hypothesis tests for means and variances, by hand and using R
+ Explore the relationship between effect size, sample size, $\alpha$, and power
+ Articulate trade-offs in study design decisions relating to effect size, sample size, $\alpha$, and power

# Exercises

# By hand or typed

1. Independent random samples of heights of adult males living in two countries yielded the following results: $n_X = 12$, $\overline{X} = 65.7$ inches, $s_X = 4$ inches and $n_Y = 15$, $\overline{Y} = 68.2$ inches, $s_Y = 3$ inches. Conduct a hypothesis test to test whether the means are equal in each population. Assume that $\sigma_X^2 = \sigma_Y^2$.

2. [HTZ 8.3-4] In May, the fill weights of six-pound boxes of laundry soap had a mean of 6.13 pounds with a standard deviation of 0.095. The goal was to decrease the standard deviation. The company decided to adjust the filling machines and then test $H_0: \sigma = 0.095$ against $H_A: \sigma < 0.095$. In June, a random sample of size $n = 20$ yielded $\bar{x} = 6.10$ and $s = 0.065$

    a. At an $\alpha = 0.05$ significance level, was the company successful?
    b. What is the approximate p-value of your test?
    
3. Let $X$ and $Y$ equal the times in days required for maturation of Guardiola seeds from narrow-leaved and broad-leaved parents, respectively. Assume that $X$ is $N(\mu_x, \sigma^2_x)$ and $Y$ is $N(\mu_y, \sigma^2_y)$ and that $X$ and $Y$ are independent. Test the hypothesis that $H_0: \sigma^2_x = \sigma^2_y$ vs. $H_A: \sigma^2_x > \sigma^2_y$ if a sample size $n_x = 13$ yielded $\bar{x} = 18.97, s_x^2 = 9.88$ and a sample of size $n_y = 9$ yielded $\bar{y} = 23.2, s_y^2 = 4.08$. Let $\alpha = 0.05$.

# Typed

## Hypothesis testing in R

For Exercises 4 - 7, you are asked to conduct hypothesis tests on the same data you constructed confidence intervals for in PS-03-04. You should copy and paste code from PS-03-04 for these exercises.

For each exercise, 

a) state the parameter of interest (in words and symbols)
b) state the null and alternative hypotheses
c) report the relevant point estimate (provide the correct notation as well as the numeric value)
d) report the test statistic, and comment on what reference distribution it is being compared to (e.g. t, standard normal, Chi-squared)
e) report the p-value
f) state your conclusion to the hypothesis test
g) report the corresponding confidence interval, and comment how it relates to your conclusion from the hypothesis test

**Note: you should use `t.test()` - you do NOT need to do these by hand.**

4. [From PS-03-04 #2] As a clue to the amount of organic waste in Lake Macatawa, researchers decide to estimate the average number of bacteria colonies in 100 milliliters of water. The number of colonies, in hundreds, for $n=30$ samples of water from the east basin yielded the following data. Use $\alpha = 0.1$ to conduct an appropriate hypothesis test for $H_0: = 70$ vs. $H_A: \mu \neq 70$.

```{r}
x <- c(93, 140, 8, 120, 3, 120, 33, 70, 91, 61, 7, 100, 19,
       98, 119, 23, 14, 94, 57, 9, 66, 53, 28, 76, 58, 9, 
       73, 49, 37, 92)
```

5. [From PS-03-04 #3] Let $X$ and $Y$ equal, respectively, the blood volumes in milliliters for a male who is a paraplegic and participates in vigorous physical activities and for a male who is able-bodied and participates in everyday, ordinary activities. Assume that $X \sim N(\mu_X, \sigma^2_X)$ and $Y \sim N(\mu_Y, \sigma_Y^2)$. The following are the $n_X = 7$ and $n_Y = 10$ observations of $X$ and $Y$. Use $\alpha = 0.05$ to conduct an appropriate hypothesis test. 

```{r}
x <- c(1612, 1352, 1456, 1222, 1560, 1456, 1924)
y <- c(1082, 1300, 1092, 1040, 910, 1248, 1092, 1040, 1092, 1288)
```

6. [From PS-03-04 #4] Students in a semester-long health-fitness program have their percentage of body fat measured at the beginning of the semester and at the end of the semester. The following measurements give these percentages for ten males in the class. Use $\alpha = 0.1$ to conduct an appropriate hypothesis test. 

```{r}
pre <- c(11.1, 19.5, 14, 8.3, 12.4, 
         7.89, 12.1, 8.3, 12.31, 10)

post <- c(9.97, 15.8, 13.02, 9.28, 11.51, 
          7.4, 10.7, 10.4, 11.4, 11.95)
```

7. [From PS-03-04 #5] Twenty-four 9th and 10th grade high school girls were put on an ultraheavy rope-jumping program. The following data give the time difference in seconds for each girl ("before program time" minus "after program time") for the 40 yard dash. Use $\alpha = 0.05$ to conduct an appropriate hypothesis test. 

```{r}
time_diff <- c(0.28, 0.01, 0.13, 0.33, -0.03, 0.07, -0.18, -0.14,
               -0.33, 0.01, 0.22, 0.29, -0.08, 0.23, 0.08, 0.04,
               -0.30, -0.08, 0.09, 0.70, 0.33, -0.34, 0.50, 0.06)
```

## Power

### Cohen's d

In many different research contexts, the parameter of interest is a mean difference. For example we are interested in the difference between the mean in the treatment group $\mu_T$ and the mean in the control group $\mu_C$. It is common practice to standardize this difference by the population standard deviation. That is, the parameter of interest is $$\delta = \frac{\mu_T - \mu_C}{\sigma}.$$ We estimate $\delta$ by $$d = \frac{\overline{Y}_T - \overline{Y}_C}{S_p}$$

Under some standard assumptions (that aren't too hard to meet in many research contexts), $d \sim N(0,1)$. The estimator $d$ above, is commonly known as Cohen's $d$, named after psychologist and statistician Jacob Cohen. 

When designing a study, you want to determine:

+ What is a practically meaningful value of $\delta$ in my research context?
+ What sample size do I need to be able to detect that effect?

Or alternatively, 

+ What is the largest sample size I can afford with my budget?
+ What is the minimal detectable effect size (MDES) with that sample size?

# Simulation exploration (in app)

Go to the following app: [https://rpsychologist.com/d3/nhst/](https://rpsychologist.com/d3/nhst/)

8. Assume that you want to detect an effect size of d = 0.1. How large of a sample do you need for the following scenarios:

    a. Type I error of 0.05 and power of 0.80
    b. Type I error of 0.01 and power of 0.80
    c. Type I error of 0.01 and power of 0.90
    
9. Let's say you want to power your test at 80% and restrict your Type I error to 0.01 and can only afford a sample size of 50. What's the smallest Cohen's d you will be able to detect under these constraints?

10. Let's say the value you determined in #9 is unrealistically large for your research context (i.e. there's no way the "treatment" will be that effective, that much better than the control), and you really need to be able to detect an effect size of 0.3. However, you don't have access to any more money - $n = 50$ is still all that you can afford. 

    a. What could you change about your design to be able to detect 0.3? Come up with at least two alternative designs. 
    b. Comment on the tradeoffs in your designs.
    
11. Suppose you want to power your study at 80% and need to detect an effect of 0.5. What happens to the necessary sample size as alpha decreases? Describe why this happens. Describe the trade-off here in practical terms.

12. Suppose a Type I error is relatively risky in your context and therefore you set $\alpha = 0.01$. You need to detect an effect of 0.5. What happens to the necessary sample size as power decreases? Describe why this happens. Describe the trade-off here in practical terms.

13. Suppose you decide to power your study at 80% and are okay with a Type I error of 5%. What happens to the necessary sample size as $d$ changes? Describe why this happens. Describe the trade-off here in practical terms. 

14. Suppose you're serving as a statistical consultant and your client  says they have budgeted for a sample size of $n = 50$. They mention that in a pilot study they estimated their treatment improved outcomes by about d = 0.1. Explain how you would go about articulating to them that this research design is not set up for success. 

## Submission

Once you have completed all Exercises and written up your results in the Template .Rmd provided, take about 5-10 minutes to respond to the reflection prompts (RP) at the top of the Template:

+ (RP1): *What were the main concepts covered in this assignment?*

+ (RP2): *What's one thing you understand better after completing these problems?*

+ (RP3): *What problems gave you the most trouble? What was difficult about them/where did you get stuck?*

Knit your document one final time, then upload your .html file to Problem Set 06 on Canvas. 

## Grading (50pts)

| Component | Points |
|:----------|:-------|
| RP1      | 5      |
| RP2      | 5    |
| RP3      | 5    |
| Formatting & Clarity | 5 |
| Exercise completion | 30 |