---
title: "Problem Set 08"
subtitle: "Week 08"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
    theme: "cosmo"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidymodels)
```


# Learning objectives


# Exercises

## By hand or typed

### Exercise 1

[HTZ: 6.5-1] Show that the residuals $Y_i - \hat{Y}_i, \ \ \ i= 1,2,...,n$ from the least squares regression model sum to 0.

### Exercise 2

[HTZ 6.5-2] In some situations where the regression model is useful, it is known that when $X = 0$ the mean of $Y$ is 0. In this case, the model $Y_i = \beta x_i + \epsilon_i,$ where $\epsilon_i$ $i = 1,2,...,n$ are independent with $\epsilon_i \sim N(0, \sigma^2)$.

a. Obtain the MLEs $\hat{\beta}$ and $\hat{\sigma}^2$ of $\beta$ and $\sigma^2$ under this model
b. Find the distribution of $\hat{\beta}$. 
c. Suppose a sample of size $n = 2$ consists of the observations $(x_1, y_1) = (1,1)$ and $(x_2, y_2) = (2,1)$. Calculate the estimates of $\beta$ and $\sigma^2$ from part (a) for these data, and plot the points and fitted line on the same graph
d. For the data in part (c), let $\hat{y_i} = \hat{\beta}x_i$ for $i = 1,2$. Calculate the residuals $y_i - \hat{y}_i$ and show that they do not sum to 0. 

### Exercise 3

Under the assumptions of the simple linear regression model, prove that a $(1 - \alpha)100\%$ confidence interval for the intercept parameter $\alpha$ is given by:

$$a \pm t_{\alpha/2}\left(\sqrt{\frac{\hat{\sigma}^2}{n-2}}\right)$$

## Typed

### Exercise 4

```{r, echo = FALSE, eval = FALSE}
library(tidyverse)

senic <- read.table("/Volumes/GoogleDrive/My Drive/STAT 511/homework/data/APPENC01.txt", quote="\"", comment.char="")

colnames(senic) <- c("ID", "length", "age", "infection_risk",
                     "ratio_culture", "ratio_xray", "num_beds",
                     "med_school", "region", "avg_census",
                     "num_nurses", "avail_facilities")

write_csv(senic, "./PROBLEM_SETS/data/senic.csv")
```

The following code reads in a dataset called `senic` which contains data on a random sample of 113 hospitals in the US. We will be exploring two variables:

+ `infection_risk` = Average estimated probability of acquiring an infection in the hospital (measured as %)
+ `length` = Average length of stay of all patients in the hospital (in days)

```{r}
senic <- read_csv("https://raw.githubusercontent.com/kgfitzgerald/APU_MATH_362/main/PROBLEM_SETS/data/senic.csv")
```

a. Fit a linear model using `infection_risk` to predict `length` of stay. Call your model `m1`
b. Write out the equation of the fitted regression model
c. Interpret the slope and intercept in context
d. Write out the hypotheses to test the slope parameter
e. Report the relevant test statistic, p-value, and your conclusion
f. Recall the test statistic is computed by $t = \frac{\hat{\beta} - \beta_0}{\sqrt{\frac{MSE}{\sum_{i=1}^n(x_i - \bar{x})^2}}}$. Use the output from `tidy(m1)` and the code below to extract the relevant numbers to compute the test statistic "by hand"

```{r, eval = FALSE}
# raw output
anova(m1$fit)

# tidy output for extracting numbers with code
tidy(anova(m1$fit))

# note the sum of squares (SS) in the denom in square root 
# of test statistic is simply the same numerator that 
# appears on the sample variance of x
# so can use the following code to compute it
n <- nrow(senic)
SS <- var(senic$infection_risk)*(n - 1)

# alternatively, to compute sum of squares (SS) "manually": 
SS_alt <- senic %>% 
  mutate(diff_sqrd = (infection_risk - mean(infection_risk))^2) %>% 
  summarize(SS = sum(diff_sqrd)) %>% 
  pull(SS)
```


```{r, echo = FALSE, eval = FALSE}
m1 <- linear_reg() %>% 
  set_engine("lm") %>% 
  fit(length ~ infection_risk, data = senic)
tidy(m1)


tidy(anova(m1$fit))


n <- nrow(senic)
SS <- var(senic$infection_risk)*(n - 1)

SS_alt <- senic %>% 
  mutate(diff_sqrd = (infection_risk - mean(infection_risk))^2) %>% 
  summarize(SS = sum(diff_sqrd)) %>% 
  pull(SS)

b <- tidy(m1)$estimate[2]
MSE <- tidy(anova(m1$fit))$meansq[2]
t <- b/sqrt(MSE/SS)
```


## Submission

Once you have completed all Exercises and written up your results in the Template .Rmd provided, take about 5-10 minutes to respond to the reflection prompts (RP) at the top of the Template:

+ (RP1): *What were the main concepts covered in this assignment?*

+ (RP2): *What's one thing you understand better after completing these problems?*

+ (RP3): *What problems gave you the most trouble? What was difficult about them/where did you get stuck?*

Knit your document one final time, then upload your .html file to Problem Set 08 on Canvas. 

## Grading (50pts)

| Component | Points |
|:----------|:-------|
| RP1      | 5      |
| RP2      | 5    |
| RP3      | 5    |
| Formatting & Clarity | 5 |
| Exercise completion | 30 |