---
title: "Gettysburg simulation"
output: 
  html_document:
    highlight: tango
    theme: "cosmo"
---

```{r message = FALSE, warning = FALSE}
library(readxl)
library(infer)
library(tidyverse)
library(broom)

set.seed(43)
gettysburg <- read_excel("./data/gettysburg.xlsx")
glimpse(gettysburg)
``` 

We can confirm that the population mean is `r round(mean(gettysburg$Length),2)`.

```{r}
gettysburg %>% 
  summarise(mu = mean(Length))
```

The plot below shows the distribution of length of words in the population. This is the *population distribution*. 

```{r}
ggplot(data = gettysburg, aes(x = Length)) +
  geom_bar() +
  geom_vline(xintercept = 4.29, color = "blue")
```

We can select a random sample of 10 words and plot the distribution of word lengths in our sample. 

```{r}
sample <- gettysburg %>% 
  sample_n(10)
sample

ggplot(data = sample, aes(x = Length)) +
  geom_bar()

sample %>% 
  summarise(avg_length = mean(Length))
```

We then select 10000 random samples of 10 words...

```{r}
sims <- rep_sample_n(gettysburg, size = 10, 
                            reps = 10000, replace = TRUE)
sims %>% print(n = 15)
```

... and compute the mean word length in each sample

```{r}
sample_means <- sims %>% 
  group_by(replicate) %>% 
  summarise(xbar = mean(Length))

sample_means %>% print(n = 15)
```

We can then plot the *sampling distribution* of the sample mean.  

```{r}
Exbar <- mean(sample_means$xbar)

ggplot(data = sample_means, aes(x = xbar)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = 4.29, color = "blue") +
  geom_vline(xintercept = Exbar, color = "green")
```

