---
title: "Hypothesis Testing"
author: "Dr. Katie Fitzgerald"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
    chalkboard: true
---

```{r child = "setup.Rmd", eval = FALSE}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
```

# Duke legend

There is a campus legend at Duke about two students who miss an exam due to excessive partying, but they tell their professor that they had a flat tire.  They realize that this sounds like a flimsy excuse, so they are pleasantly surprised when the professor accepts their explanation and offers a make-up exam on the following morning.  

When they arrive for the make-up exam, they are sent to two separate rooms.  They find question 1, worth 5 points, to be quite straight-forward.  Then they turn the page to find question 2, worth 95 points: *Which tire was it?*

---

# Do people disproportionately choose the right front tire?

In a previous semester, 17 out of 44 students chose the right front tire. 

1. Identify the observational units and variable. Is the variable categorical or numeric?

--

2. Calculate the proportion of these students who selected the right front tire.  Is this a parameter or a statistic?  Explain.

--

3. Write a sentence describing the parameter of interest

--

4. State (in words) the null and alternative hypotheses to be tested.

--

5. How might we determine if this constitutes sufficient evidence that people disproportionately choose the right front tire? 

---

# Simulation logic

Write out how you could use simulations to determine how (un)likely the observed result it.

---

# Simulation applet

[https://www.rossmanchance.com/applets/2021/oneprop/OneProp.htm](https://www.rossmanchance.com/applets/2021/oneprop/OneProp.htm)

---

# Using test-statistic/critical region

1. Calculate and interpret the value of the test statistic. 

2. How is this test statistic distributed?

3. Determine an appropriate critical region if you want your probability of making a Type I error to be no larger than 1%. 

4. Draw a picture to illustrate your critical region and your test statistic.

5. Summarize your conclusion, and provide justification based on the test statistic

---

# "Which tire?" example continued

At a recent workshop for college professors, 8 of 36 workshop participants selected the right front tire.

1. Explain why it’s not necessary to carry out calculations for a hypothesis test of whether the sample data provide strong evidence that people select the right front tire more often than would be expected by random chance.

2. Without performing an analysis, what can you say about how the critical value/test statistic would turn out? 

3. Based on this sample result, would you reject the null hypothesis?  Explain.

4. Based on this sample result, would you accept the null hypothesis?  Explain.

---

# Modus Tollens

+ If P then Q
+ Not Q
+ Therefore not P

--

Example: the Constitution of the United States stipulates that if a person is eligible to be elected President in the year 2024 (call this P), then that person must have been born in the U.S. (call this Q).

*Queen Elizabeth was born outside the U.S. What can we know about whether she is eligible to be elected president in 2024?*

--

*Zendaya was born inside the U.S. What can we know about whether she is eligible to be elected president in 2024?*

---

# For right tire... $\hat{p} = 8/36 = 0.22$

P = null hypothesis that people choose randomly (no preference for right tire)

--

Q = test statistic is less than `pnorm(.99) = 2.326`. Or equivalently, $\hat{p} < .305$

--

Here we have an *absence of evidence* that people disproportionately prefer right tire, but...

--

ABSENCE OF EVIDENCE DOES NOT CONSTITUTE EVIDENCE OF ABSENCE

--

*Many different hypotheses are consistent with the observed data, so it’s not appropriate to accept any ONE of these hypotheses.*