---
title: "Exam 2"
subtitle: "MATH 362"
date: "due Thursday, November 17, 11:59pm"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
    theme: "paper"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warning = FALSE}
library(infer)
library(kableExtra)
library(tidyverse)
```

```{css, echo = FALSE}
/* Whole document: */
body{
  font-size: 16pt;
}
/* Headers */
h1{
  font-size: 24pt;
}
```

For any hypothesis tests on this exam, use $\alpha = 0.01$. 

## Problem 1

Let $Y_1, Y_2, ..., Y_n$ be independent random variables from a distribution with mean $\mu$ and standard deviation $\sigma$. Suppose we are interested in estimating $\mu^2$. Let $\bar{Y} = \frac{\sum_{i = 1}^nY_i}{n}$ and $S^2 = \frac{\sum_{i =1 }^n(Y_i - \bar{Y})^2}{n-1}$ be the usual sample mean and sample variance estimators. Find a constant $k$ (possibly depending on $n$ but not on $\mu$ or $\sigma$) such that $\overline{Y}^2 -kS^2$ is an unbiased estimator of $\mu^2$.

## Problem 2

Let $Y_1, Y_2, ..., Y_n$ be independent random variables with pdf 
$$f(y;\theta) = \theta(1-\theta)^y, \ \ \ y = 0,1, \dots; 0 < \theta < 1$$
Find the MLE of $\theta$.

## Problem 3

Let $Y_1, Y_2, ..., Y_n$ denote independent random variables each with pdf $$f(y) = \lambda exp(-\lambda y), \ \ \ y>0;$$

This is an exponential distribution with parameter $\lambda$. Let $\overline{Y} = \sum_{i = 1}^n \frac{Y_i}{n}$ and consider the random variable $$W = exp(-1/\overline{Y}).$$

The distribution of $W$ is not known.

a. Use simulations to estimate $E(W)$ when $n = 25$ and $\lambda = 3$. Some starter code is provided below.

```{r, eval = FALSE}
#initialize empty vector to store sample means
ybar <- c()

#for loop to generate and store a large number of 
#random ybar values
for(i in 1:___){
  #generate random sample of size n 
  #from exponential distribution with rate parameter lambda
  sample_i <- rexp(n = ___, rate = ___)
  #compute sample mean for ith sample
  ybar[i] <- mean(sample_i)
}

#compute W

#estimate expected value of W

```

b. Although the distribution of $W$ is unknown, a linear approximation method can be used to derive the following approximate formula $$E(W) \approx exp(-\lambda)\left[1 + \frac{\lambda^2 - 2\lambda}{2}\frac{1}{n}\right]$$ How does the approximation obtained from this formula compare to the approximation from your simulations in part a? 
c. Compare the approximations from the two methods (i.e. simulations & the above formula) for $n = 5$ and again for $n = 250$. (*Note: this will involve running your simulations two more times*). Comment on your findings. 

```{r, echo = FALSE, eval = FALSE}
ybar <- c()
for(i in 1:10000){
  sample_i <- rexp(250, rate = 3)
  ybar[i] <- mean(sample_i)
}
data <- data.frame(ybar, replicate = seq(1:length(ybar)))

data <- data %>% 
  mutate(W = exp(-1/ybar))

mean(data$W)
lambda <- 3
n <- 250
exp(-lambda)*(1 + (lambda^2 - 2*lambda)/2*1/n)
```

## Problem 4

The director of admissions of a small college selected 120 students at random from the new freshman class in a study to determine whether a student's GPA at the end of freshman year can be predicted from their ACT score. The data from this study can be read in using the following code. 

```{r, message = FALSE, warning = FALSE}
data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vQEj0DlMoCdeVSI9qEXv9GhFeemvXddhu7pjgJRAx-t7ljEpGilLLA3kEPrJO3_52ZIi_bPnv4ItZ2O/pub?gid=0&single=true&output=csv")
```

a. Fit a linear regression model to these data. 
b. Write out the estimated regression model.
c. Interpret the slope and intercept in context.
d. Write out the hypotheses to test the slope parameter. 
e. Report the relevant test statistic and the critical value it should be compared to.
f. State your conclusion in context.

## Problem 5

A rental car company wants to investigate whether the type of car rented affects the length of the rental period. An experiment is run for one week at a particular location, and 12 rental contracts are selected at random for each car type. The data are provided below.

```{r}
data <- data.frame(length = c(3,5,3,7,6,5,3,2,1,6,5,3,
                              1,3,4,7,5,6,3,2,1,7,6,1,
                              4,1,3,5,7,1,2,4,2,7,2,4,
                              3,5,7,5,10,3,4,7,2,7,9,7),
                   type = c(rep("subcompact", 12),
                            rep("compact", 12),
                            rep("midsize", 12),
                            rep("full size", 12)))
```

```{r, echo = FALSE, eval = FALSE}
tidy(aov(length ~ type, data))
```

a. State the hypotheses to test whether there is evidence to support a claim that the type of car rented affects the length of the rental contract. 
b. Conduct an ANOVA on these data. 
c. Report the following values for this analysis (according to the PSU notation):
    i. $m$
    ii. $n$
    iii. degrees of freedom for the $SSE$
    iv. $SSB$ (sometimes called $SST$; B = between or T = treatment)
    v. $MSE$
    vi. $F$
    vii. p-value
d. What can be concluded from this analysis? Justify your conclusion. 