---
title: "Exam 2 Review"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
    theme: "paper"
---

## Exam Info

+ 5 problems, 20 points each
    + 1 biased/unbiased estimator problem
    + 1 MLE problem
    + 1 simulation problem
    + 1 regression problem
    + 1 ANOVA problem

## Practice Problems

### (Un)biased estimators

Let $Y_1, Y_2$ be independent Poisson random variables with mean $\lambda$ and let $$\hat{\lambda}_1 = \frac{Y_1 + Y_2}{2}$$
and 
$$\hat{\lambda}_2 = \frac{(Y_1 - Y_2)^2}{2}$$
denote two different estimators of $\lambda$. Show that $\hat{\lambda}_1$ and $\hat{\lambda}_2$ are both unbiased estimators of $\lambda$. 


### Simulation

Use simulations to demonstrate that the two above estimators are unbiased. Use $\lambda = 5$. Some starter code is provided.

```{r, eval = FALSE}
Y1 <- rpois(___, ___)
Y2 <- rpois(___, ___)

# compute lambda1_hat

# compute lambda2_hat

# estimate expected value of lambda1_hat

# estimate expected value of lambda2_hat


```

```{r, eval = FALSE, echo = FALSE}
Y1 <- rpois(10000, 5)
Y2 <- rpois(10000, 5)

# compute lambda1_hat
lambda1_hat <- (Y1 + Y2)/2

# compute lambda2_hat
lambda2_hat <- (Y1 - Y2)^2/2

# estimate expected value of lambda1_hat
mean(lambda1_hat)

# estimate expected value of lambda2_hat
mean(lambda2_hat)

```

Use simulations to estimate the expected value of the sum of two Poisson random variables $W = Y_1 + Y_2$. 

```{r, echo = FALSE, eval = FALSE}
mean(lambda1_hat + lambda2_hat)
```


### MLE

Let $Y_1, Y_2, ..., Y_n$ be independent random variables with pdf $$\lambda exp\{-\lambda y\}, \ \ \ y > 0,$$
where $\lambda > 0$ is an unknown parameter. Find the MLE of $\lambda$.



### Regression

The `mtcars` dataset has data on design and performance of 32 cars from the 1974 *Motor Trend* US magazine. 

a. Fit a regression model using weight (1000 lbs) to predict miles per gallon.   
b. Write out the regression equation
c. Conduct a hypothesis test for the slope parameter. Make sure to state the hypotheses, the test statistic, the p-value, and your conclusion. 
d. What distribution is the test statistic being compared to? What is the appropriate critical value? 

```{r, eval = FALSE, echo = FALSE}
library(broom)
library(tidyverse)
fit <- lm(mpg ~ wt, data = mtcars)
tidy(fit)
```


### ANOVA

The following output was obtained from a computer program that performed a two-factor ANOVA on a factorial experiment.

| Source      | DF                 | SS                 | MS                 | F                  | p-value            |
|------------|------------|------------|------------|------------|------------|
| Factor      | 3                 | 38.25              | \_\_\_\_\_\_\_\_\_ | \_\_\_\_\_\_\_\_\_ | \_\_\_\_\_\_\_\_\_ |
| Error       | \_\_\_\_\_\_\_\_\_| \_\_\_\_\_\_\_\_\_  | \_\_\_\_\_\_\_\_\_|                    |                    |
| Total       | 19                 | 196.04            |                    |                    |                    |

: One-way ANOVA: y versus A

a. Fill in the missing blanks
b. Write out the hypothesis that is being tested
c. What is the appropriate critical value to compare the test statistic to? 
d. What is the appropriate conclusion from this analysis?
