---
title: "Week 06 Day 01"
subtitle: "Exam 1 Review"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
    theme: "cosmo"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Exam Info

+ 1 hypothesis test question (can use `t.test()`)
+ 1 Type I / Type II error problem
+ 1 simulation problem - most code provided, need to understand conceptually
+ 5 True/False + justification of your response
+ 5 "State a research question for this parameter of interest"

## Hypothesis test example

Research question: are Lyft and Uber prices the same on average? We have a random sample of 10 rides from each company. 

```{r, echo = FALSE}
set.seed(437)
lyft <- rnorm(10, 25, 5)
uber <- rnorm(10, 30, 5)
```

```{r}
lyft <- c(26.18, 25.76, 14.38, 33.58, 22.27,
          20.0, 21.86, 28.5, 24.92, 21.74)

uber <- c(23.0, 29.56, 28.92, 37.26, 29.33,
          34.9, 23.98, 29.6, 29.1, 30.75)

t.test(lyft, uber)
```

+ What does each value in the output mean?
+ What arguments can we add/change in `t.test()`? 
+ Remember to report the notation as well as the actual number of the point estimate

## Problem Sets Revisited

### PS 02 Exercise 4

Suppose that a region has two hospitals. Hospital A has about 10 births per day, and hospital B has about 50 births per day. About 50% of all babies are boys, but the percentage who are boys varies at each hospital from day to day. Over the course of a year, which hospital will have more days on which 60% or more of the births are boys – A, B, or negligible difference between A and B?

### PS 02 Exercise 6

Why are the following statements INCORRECT:

A. We are 95% confident that a thriller film in the IMDb list runs for between 94.5 and 108.9 minutes.

F. We are 95% confident that the mean runlength of a thriller film in this sample from the IMDb list was between 94.5 and 108.9 minutes.

G. We are 95% confident that the mean runlength of a thriller film in a new random sample from the IMDb list would be between 94.5 and 108.9 minutes.

### PS 02 Exercise 8

Why are the following not appropriate research questions when the parameter of interest is a mean?

+ How many people _____?
+ What is the average number of people in this city who _____?

### PS 03-04 Exercise 4 / 5

What notation do we use when dealing with paired means?

+ E.g. pre-post measurements with fitness program
+ E.g. race time differences (before time - after time)

### PS 03-04 Exercise 11

What's the difference between a sample distribution and sampling distribution?

### PS 03-04 Exercise 15 - 20

Is $\hat{p}$ an unbiased estimate of $p$? 

+ [https://kgfitzgerald.github.io/APU_MATH_362/PROBLEM_SETS/SOLUTIONS/PS-03-04-SOLUTIONS.html#Exercise_15](https://kgfitzgerald.github.io/APU_MATH_362/PROBLEM_SETS/SOLUTIONS/PS-03-04-SOLUTIONS.html#Exercise_15)

What went wrong? 

Do we prefer coverage probabilities that are above or below the nominal level? 

### PS 05 

How often will $T < 1.96$ when $H_0$ is true? $T > 0$?

How often will $p-value < 0.01$ when $H_0$ is true? $p-value < 0.23$?

## Your Questions? 

## Rest of class

+ Explain PS 05 Exercise 3 to each other. **Also articulate (in detail) how you could use simulations to demonstrate Type I and Type II error.**  

+ Write a research question for each parameter of interest we've covered so far. Exchange with your group members and give feedback.

+ Take turns explaining/teaching Knowledge Check questions to each other

+ Together, come up with a list of 5 - 10 true false questions that could be asked to evaluate knowledge of material covered so far