---
title: "Week 02 Day 02"
subtitle: ""
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
    theme: "cosmo"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


<!-- ## Reading Annotation Q/A -->

<!-- 1. Should we always assume random data is independent? -->

<!-- 2. What do we do when $\sigma_X^2 \neq \sigma_Y^2$? -->

<!-- 3. "[Paired t-test] would only work if sample sizes m and n are equal correct? Or else some of the observations would be left unpaired? But I can't help but wonder: if this was like a before and after study of the same subjects, but for whatever reason, data was not obtained "afterwards" for some of the subjects, what do we do then? Do we just ignore those observations and use this method, limiting our sample size only to those we have data for before and after?" -->


## Knowledge Check

For Questions 1-5, use the following scenario. Let $X_1, X_2, \dots X_n \sim N(\mu_X, \sigma^2)$ and $Y_1, Y_2, \dots Y_n \sim N(\mu_Y, \sigma^2)$ be independent random samples.

1. How is $\overline{X} - \overline{Y}$ distributed?
1. How is $\frac{(\overline{X} - \overline{Y}) - (\mu_X - \mu_Y)}{\sqrt{\frac{\sigma^2}{n} + \frac{\sigma^2}{m}}}$ distributed?
1. How are $\frac{(n-1)S_x^2}{\sigma^2}$  and $\frac{(m-1)S_y^2}{\sigma^2}$ distributed?
1. How is $\frac{(n-1)S_x^2}{\sigma^2} + \frac{(m-1)S_y^2}{\sigma^2}$ distributed?
1. If $Z \sim N(0,1)$ and $U \sim \chi^2_{df}$, how is $\frac{Z}{\sqrt{U/df}}$ distributed? 

1. When should you use Welch's t-interval? 
1. When should you use a paired t-interval?

## Simulation exploration

For this section, use the [Rossman/Chance applet](https://www.rossmanchance.com/applets/2021/confsim/ConfSim.html) for Confidence Intervals. To begin, assume that the population has a normal distribution with mean 50 and standard deviation 10.  Use a very small sample size of 5, a confidence level of 95%, and simulate selecting 500 random samples from the population. Select "z with s" for the Method dropdown. 

1. What problem with the “z with s” confidence interval procedure does this simulation analysis reveal?

1. In order to solve this problem, do we need the intervals to get a bit narrower or wider?

1. Which of the four terms in the formula $\bar{x} \pm z_{\alpha/2}\frac{s}{\sqrt{n}}$ (i.e. $\bar{x}, z_{\alpha/2}, s$ or $n$) can we alter to produce a narrower/wider interval?  In other words, which one does not depend on the data?

1. Do we want to use a larger or smaller multiplier than $z_{\alpha/2}$?

1. Use the `qt()` function in R to find the appropriate t-multiplier. How does this multiplier compare to the value of $z_{\alpha/2}$ for 95% confidence? 

```{r, echo = FALSE, eval = FALSE}
qt(0.975, df = 4)
```

6. Change the Method to "t" in the applet. How do these intervals compare to the previous ones? Do more of them succeed in capturing the parameter value?

7. What happens with larger sample sizes? Conduct more simulations with larger sample sizes, and compare "z with s" and "t" for each. What do you notice? 

8. Does t* work equally well with other confidence levels?

9. Why do the widths of these intervals vary from sample to sample? 

10. Change the Distribution to Uniform. Choose an upper and lower limit so that the mean is 50. Run 500 simulations with a sample size of 5. What percent capture the mean using "t" and "z with s"? Do the same for sample size of 20. What do these 4 percentages suggest? 

11. Explore trends when the data come from an Exponential distribution, using samples sizes of 5, 20, 40, 100. What do you notice? What conclusions can you draw? 