---
title: "Week 02 Day 01"
subtite: "Knowledge Check"
output: 
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Knowledge Check

1. How is $\overline{X}$ distributed? What conditions does this depend on? 
1. How is $\frac{\overline{X} - \mu}{\sigma/\sqrt{n}}$ distributed? What symbol do we use for this expression?

1. Draw a picture that illustrates what $z_{\alpha/2}$ stands for.

1. Why is it incorrect to say "The probability $\mu$ falls between L (lower bound) and U (upper bound) is $1-\alpha$"?

1. What percentage of 80% confidence intervals will "fail"? What does it mean to fail? 

1. What is the difference between $\overline{X}$ and $\mu$? 

1. What does $s$ mean? How does it differ from $\sigma$? 

1. When should we use the t-distribution? Why? 

1. How is $\frac{\overline{X} - \mu}{s/\sqrt{n}}$ distributed? What symbol do we use for this expression?

1. How many *parameters* does the normal distribution have? What are they called/how are they denoted? What about the t-distribution? 

1. What is the difference between a point estimate and a parameter? Give three examples of point estimates and their corresponding parameters. 

1. What is the difference between standard deviation, standard error, and margin of error? How are they related to one another? 

1. Explain the notation $t_{\alpha/2, n-1}$. Draw a picture to illustrate it. 

## Simulation exploration

For this section, use the [Rossman/Chance applet](https://www.rossmanchance.com/applets/2021/confsim/ConfSim.html) for Confidence Intervals. To begin, assume that the population has a normal distribution with mean 50 and standard deviation 10.  Use a very small sample size of 5, a confidence level of 95%, and simulate selecting 500 random samples from the population. Select "z with s" for the Method dropdown. 

1. What problem with the “z with s” confidence interval procedure does this simulation analysis reveal?

1. In order to solve this problem, do we need the intervals to get a bit narrower or wider?

1. Which of the four terms in the formula – x-bar, z*, s, or n – can we alter to produce a narrower/wider interval?  In other words, which one does not depend on the data?

1. Do we want to use a larger or smaller multiplier than z*?

1. Use the `qt()` function in R to find the appropriate t-multiplier. How does this multiplier compare to the value of $z^*$ for 95% confidence? 

```{r, echo = FALSE, eval = FALSE}
qt(0.975, df = 4)
```

6. How can we tell whether this t* adjustment is the right amount to produce 95% confidence?

1. Change the Method to "t" in the applet. How do these intervals compare to the previous ones? Do more of them succeed in capturing the parameter value?

1. What happens with larger sample sizes? Conduct more simulations with larger sample sizes, and compare "z with s" and "t" for each. What do you notice? 

1. Does t* work equally well with other confidence levels?

1. Why do the widths of these intervals vary from sample to sample? 

1. Change the Distribution to Uniform. Choose an upper and lower limit so that the mean is 50. Run 500 simulations with a sample size of 5. What percent capture the mean using "t" and "z with s"? Do the same for sample size of 20. What do these 4 percentages suggest? 

1. Explore trends when the data come from an Exponential distribution, using samples sizes of 5, 20, 40, 100. What do you notice? What conclusions can you draw? 