---
title: "Week 03 Day 02"
subtitle: "Knowledge Check"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
    theme: "cosmo"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Announcements

+ PS 03-04 due Saturday
+ W03 D02 and W04 D01 Reading Annotations due Monday

## Knowlege Check

1. What  formula do we use to compute the sample size needed for estimating $\mu$?

$$n \approx \frac{z_{\alpha/2}s^2}{\epsilon^2}$$

2. What is the hardest part about using the above formula? Why?  

<br/>
<br/>
<br/>

3. Comment on what each symbol in the formula in #1 means and how you obtain values to plug in.

<br/>
<br/>
<br/>

4. When estimating a mean, will increasing your level of confidence require more or less necessary sample size? Why?

<br/>
<br/>
<br/>


5. When estimating a mean, will increasing your margin of error require more or less necessary sample size? Why?

<br/>
<br/>
<br/>

6. What  formula do you use to compute the sample size needed for estimating $p$?  

$$n \approx \frac{z_{\alpha/2}\hat{p}(1-\hat{p})}{\epsilon^2}$$

7. What is the hardest part about using the above formula? Why?

<br/>
<br/>
<br/>

8. Comment on what each symbol in the formula in #6 means and how you obtain values to plug in. 

<br/>
<br/>
<br/>

9. Comment on how and why the formulas in #1 and #6 relate to one another.

<br/>
<br/>
<br/>

10. When estimating $p$, what happens to the necessary sample size as $\hat{p}$ changes? 

<br/>
<br/>
<br/>

11. Small sample correction for finite populations:

$$\frac{N-n}{N-1}$$

$$SE(\hat{p}) = \sqrt{\frac{\hat{p}(1-\hat{p})}{n}\frac{N-n}{n-1}}$$
This correction factor goes to 1 pretty quickly:

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
N <- seq(from = 500, to = 1000000, by = 1)
n <- 100

data <- data.frame(N = N,
                   n = 100,
                   correction = (N-n)/(N-1))

ggplot(data, aes(x = N, y = correction)) +
  geom_line()

ggplot(data, aes(x = N, y = correction)) +
  geom_line() +
  xlim(0, 50000)
```




